3D Regression Heat Map Analysis of Population Study Data

### 2015-06-24 draft.tex
added Bernhards second revision; added Bibs

%\com{New Solution}.
%%
We propose an analysis approach of epidemiological data sets by incorporating all features in an exhaustive regression-based analysis.
%%
This approach combines all \emph{independent features} w.r.t. a \emph{target feature}.
%%
It provides a visualization that reveals insights into the data by highlighting relationships.
%%
The \emph{3D Regression Heat Map}, a novel 3D visual encoding, acts as an overview of the whole data set.
%%
It shows all combinations of two to three independent features with a specific target disease.
%%
%A 3D visualization of all combinations of two to three independent features towards a target acts as an overview of the whole data set, the \emph{3D Regression Heat Map}.
%%
Slicing through the \emph{3D Regression Heat Map} allows for the detailed analysis of the underlying relationships.
%%
Expert knowledge about disease-specific hypotheses can be included into the analysis by adjusting the regression model formulas.
%%
Furthermore, the influences of features can be assessed using a difference view comparing different calculation results.
%%
%\com{Validation \& Results}.
%%
We applied our \emph{3D Regression Heat Map} method to a hepatic steatosis data set to reproduce results from a data mining-driven analysis.
%%
A qualitative analysis was conducted on a breast fat data set.
%%
We were able to derive new hypotheses about relations between breast fat density and breast lesions with breast cancer.
%%
%\com{Implications}.
%%
With the \emph{3D Regression Heat Map}, we present a visual overview of epidemiological data that allows for the first time an interactive regression-based analysis of large feature sets with respect to a disease.
} % end of abstract

@@

Testing features for associations with diseases using regression models is one of the most important epidemiological tools.
%%
Using regression analysis to assess the statistical resilience of a hypothesis rarely involves \emph{more than three features} due to the higher dimensional problem and the required subject count.
%%
Due to the amount of data and only limited overview visualizations, possible correlations may be missed.
%%
Explorative analyses and overview visualizations of the data set as presented in prior work \cite{Klemm2014VIS} are not tailored to a specific target feature.
%%
They mostly highlight correlations between features, which are known to the domain expert (e.g., correlation between body size and spine shape).
%%
We incorporate the regression analysis, which is familiar to the domain experts, into overview visualizations to support a hypothesis-free analysis or an analysis w.r.t. a specific disease or hypothesis.
%%
Therefore we provide template regression formulas, which are applied to all potential feature combinations.
%%
Since the notation is familiar to epidemiologists, they can rapidly include their domain knowledge into the analysis process.
%%
Difference views between regression formulas allow to assess the influences of individual features on the process.
%%
Our contributions are:

@@

We described an Interactive Visual Analysis approach for image-centric cohort study data, which connects to the feature listing step \cite{Klemm2014VIS}.
%%
The methods aim to derive hypotheses through analysis of the data and observations about previously unknown feature correlations.
%%
Employing \emph{hypothesis generation} requires overview visualizations of feature correlations, which are not supported by standard statistical processors.
%%
In this work, we focus on a similar approach to derive insight and even new hypotheses through the data rather than only using it for a confirmatory analysis.
%%
\subsection{Epidemiological Data} \label{sec:EpidemiologicalData}
%%
There are many data acquisition modalities available to epidemiologists. %, which are suitable for different analysis types.
%%
We focus on population studies, which impose the largest data sets and yield a highly heterogenous and incomplete information space.
%%
Subject data are collected either regarding specific diseases or with the widest range possible.
%%
The latter allows the data set to be assessed w.r.t. different diseases.
%%
The feature space comprises information about lifestyle, somatometric features, medical parameters, genetic data as well as medical images.
%%
These features are gathered through different modalities, such as questionnaires, medical examinations or laboratory analyses.
%%
Many features are sparse, such as follow-up questions about a medication or treatment of a certain disease.
%%
Other features are exclusive for a sub-group, such as women-specific questions, e.g., number of born children or period status.
%%
Medical status features or lifestyle factors are primarily of \emph{dichotomous} (binary) type.
%%
Continuous data are often discretized (e.g. 10 year steps for age) to equalize the feature types and to simplify the method selection.
%%
However, this should be avoided, since it reduces the information space.
%%
It also introduces a new information bias, as assumptions are modeled through the discretization.
%%

Medical image data are also acquired and analyzed in modern population studies.
%%
We incorporate image-derived features, but do not focus on analyzing image data.
%%
More discussions of population study data types and characteristics can be found in the works of Preim et al.~\cite{Preim2014} and Toennies et al.~\cite{Toennies2015}.
%%

Features influencing the exposure as well as the outcome of an analysis are called \emph{confounders} and have to be specially treated.
%%
The analysis model has to be adjusted by normalizing all included features w.r.t. the confounder.
%%
\emph{Age} is included as confounder in almost any epidemiological analysis, since most diseases (such as different cancer types) are more likely for older subjects.
%%
It also influences the general body condition and thereby almost all features acquired through population studies.
%%
Another important confounder is \emph{gender}.
%%
Confounders have to be selected by epidemiologists specific to the investigated condition.
%%
\paragraph{The Study of Health in Pomerania (SHIP).}
The \texttt{SHIP}, located in Northern Germany, aims to characterize health and disease in the widest range possible \cite{Volzke2011}.
%%
%It does not focus on a specific disease, making the data set open for many diseases.
%%
Unique for the \texttt{SHIP} is the acquisition of medical image data per participant.
%%
A second cohort, \texttt{SHIP-TREND} was started in 2012.
%%
Data for both cohorts are examined in a 5-year time span.
%%
New parameters are added in each iteration, extending the range of investigated diseases.
%%
%For the latest two acquisitions (\texttt{SHIP-2} and \texttt{SHIP-TREND-0}), MRI scans are included into the cohort \cite{Hegenscheid2009, Ivanovska2014}.
%%
Most examinations occur in all stages and are performed according to the same instructions to enable comparisons over the different stages of the study.

\subsection{Regression Analysis} \label{sec:RegressionAnalysis}
%%
Regression analysis is the most important statistical tool when analyzing epidemiological data and impose the basis of this work.
%%
A regression analysis assesses the influence of one or more (\emph{independent}) features to one target (\emph{dependent}) feature.
%%
The regression model yields a function describing the target feature by weighting the independent features.
%%
Different metrics, such as the weightings itself and associated $p$ values, describe the resulting function (the \emph{model}).
%%
\add{
%%
The \emph{Akaike Information Criterion} (AIC) metric estimates the quality of a model and is suited for comparing models.
%%
A small AIC value indicates a higher quality model.
%%
}
$R^2$ values describe the quality of the fit; in other words how well the dependent features describe the target feature.
%%
The value ranges between $[0, 1]$, where 1 encodes a perfect fit.
%%

@@

\paragraph{Logistic Regression for Dichotomous Target.} Logistic regression implies a dichotomous target variable.
%%
The target is described by fitting a logistic function.
%%
Logistic models, as opposed to linear models, do not allow for extracting an $R^2$ quality of fit value.
%%
Therefore, pseudo-$R^2$ values are extracted, such as the \emph{Nagelkerke $R^2$}, which mimics the behavior of the $R^2$.
%%
\emph{Nagelkerke $R^2$} behaves different than $R^2$ values extracted from the linear regression model.
%%
Comparisons have to be handled with care.
%%
\section{Prior and Related Work}
%% https://en.wikipedia.org/wiki/Exploratory_data_analysis#cite_note-Tukey1977-4
Tukey already stated in 1977 that data are too often analyzed solely using confirmatory data analysis \cite{Tukey}.
%%
He emphasized the need to use data to \emph{derive} hypotheses, which can then be tested again.
%%
In this section we present prior and related work trying to achieve this goal.
%%
%The work of M\"{u}hlbacher et al.~\cite{Muehlbacher}, Piringer et al.~\cite{Piringer}, Chan et al.~\cite{Chan} and Zhang et al.~\cite{Zhang2014} are closest related to ours regarding different aspects.
\paragraph{Parameter space analysis using regression models.}
%%
\magenta{
Sedlmair et al.~\cite{Sedlmair} present a conceptual taxonomy of parameter space analysis.
%%
%It includes a abstract application domain independent data flow model, navigation strategies for exploring parameter spaces as well a characterization of analysis tasks.
%%
%Using the taxonomy, approaches can be classified to find works with similar problems and derive new design ideas and evaluate ideas.
%The taxonomy allows to classify approaches to find works with similar problems to derive new design ideas and evaluate ideas.
%%
Based on their input parameter taxonomy, we are visualizing \emph{model parameters} based on \emph{environmental parameters} in a \emph{global-to-local} navigation strategy.
%%
We solve a \emph{fitting} task by aiming to find models well suited for describing the input data.
}

\magenta{
%% A Partition-based Framework for Building and Validating Regression Models
The approach of M\"{u}hlbacher et al.~\cite{Muehlbacher} is closest to ours.
%%
They provide a framework for qualitative analyses of relationships and ranking features for numerical target features with regression models.
%%
%Correlations between features with the target yield a ordered matrix-plot, where feature combinations are used to depict models of interest.
%A view for single-relationships with the target and single features as well as a matrix-plot for feature combinations are used to depict models of interest.
%%
%The visualization allows to assess different model complexities.
%%
Existing regression models can be validated and compared using 3D views and 2D slice-views.
%%
%We employ the same steps by preprocessing features using correlations with the target and providing an overview over combinations.
%%
M\"{u}hlbacher et al. focus on a smaller number of features, which can be assessed in more detail, yielding a plot-matrix view, while we cover more features by abstracting the models.
%%
%We also take categorical targets and regression formula inputs into account.
%%
}

@@

\paragraph{Visual analysis of epidemiological data.}
%%
Zhang et al.~\cite{Zhang2014, Gotz2014} is closest to our work regarding the application of visual analysis on cohort study data.
%%
They present \emph{Cohort Analysis via Visual Analytics} (\emph{CAVA}), a framework that distinguishes three major elements of cohort study data analysis: \emph{cohort data} (and its manipulation using operations), \emph{views} and \emph{analytics}.
%%
They use the system to find longitudinal pathways for diseases on the basis of health records.
%%
%Sub-groups are automatically created by incorporating \emph{analytics} and through selection in expert-guided \emph{views}.
%%
%They do not go into the detail w.r.t the analytics used, they seem to not incorporate confounder analysis (other than dividing groups by gender) and the user study lacks concrete goals.
%%
%Their data differs strongly from ours, having multiple moments for each subject, yielding a approach focused on finding pathways of interest.
%%
We apply their requirements as guidelines, which incorporate \emph{flexible} and \emph{iterative analysis}. %, meaning that the system has to adapt to different requirements in different analysis steps.
%%
%We implement this by supporting both hypothesis-based and hypothesis-free analysis.
%%
Angelelli et al.~\cite{Angelelli} visualize image-derived an non-image data using cube data structures with focus on comparison and knowledge extraction.
%%
%Their proposed engine represents multiple heterogenous cohort study data sets as normalized data cubes.
%%
%Thus, data redundancy is avoided and runtime measure aggregation can be carried out when measures of the different cubes have to be combined.
%%
%This allows for saving memory space.
%%
%Their presented aggregation functions allow for a fast data calculation to map it using different linked visualizations, such as scatter plots, bar charts or histograms.
%%
They use \emph{Pearsson's} $r$ to characterize relationships with target features and employ list views and scatter plots to visualize and rank them.
%%
%We use regression models and associated $R^2$ values to depict relationships between a target feature and multiple independent features instead of one and employ the information into one global visualization.
%%
In contrast, we aim for a fast large scale correlation analysis incorporating many features to derive insights.
%%
%\paragraph{Visual Analysis in Public Health.}
%%--
% Shneiderman et al.~\cite{Shneiderman2013} highlight the necessity of interactive visualizations in healthcare data. %, e.g. using \emph{Health 2.0}, which is an umbrella for web-based health-informatics frameworks.
% %%
% %They divide Health 2.0 in three domains, \emph{Personal health information}, \emph{clinical health information}, and \emph{public health information}.
% %%
% %We focus on the latter.
% %%
% %Shneiderman and colleagues also present challenges,
% The presented challenges in this application require the systems to be sufficiently comprehensible and intuitive to fit into the short time span that clinicians have for analyzing data.
% %%
% Furthermore, the systems should provide comparative relationship visualization, characterization of similarity and presenting risks.
%%
%We use the challenges as design guidelines for our web-based prototype, focusing on functional clean designs to allow fast insight into relationships towards diseases.
%%

@@

%% Steenwijk: Integrated Visual Analysis for Heterogeneous Datasets in Cohort Studies
Steenwijk et al.~\cite{Steenwijk} focus on \emph{hypothesis-free} exploration of cohort data.
%%
They employ a framework consisting of feature extraction and visualization to derive dependencies between image- and non-image features.
%%
They incorporate linked views using scatter plots, bar charts, parallel coordinates as well as time plots to display and brush the data.
%%
%Medical image data can be incorporated to investigate potentially interesting individuals.
%%
%%We distinguish to this and other linked-view analysis based approaches by showing abstracted information derived from a preceding analyze-first step to highlight relationships rather than displaying the actual data using various information visualization techniques.
%%
The techniques are useful for selected features or a data set with a small number of features due to their increase in complexity with every additional parameter.
%%
They are therefore suitable for \emph{small} epidemiological data sets.
%%
\magenta{
Maries et al.~\cite{GRACE} propose \texttt{GRACE}, a framework for visually exploring correlation of features with spatial and non-spatial geriatric data.
%%
Sparse Partial Least Squares Regression and Tikhonov Regularization are used to produce predictor subsets and quantify correlation strengths.
%%
A multi-view system linked spatial with non-spatial data and correlation coefficients are visualized using Kiviat diagrams (similar to star plots), correlations are plotted using p-values in the other views.
%%
%Dendograms are used to select different brain regions.
Instead of applying several linked views, we focus on one view of abstracted data.
}

Generalized Pairs plots (\texttt{GPLOM'S}) extend the concept of scatter plot matrices by the pairwise depiction of heterogeneous data using type-combination-dependent visualizations \cite{GPLOMS, Francois2013}.
%%
Dai et al.~\cite{Dai2005} incorporate a \texttt{GPLOM}-like visualization using choropleth maps mapping spatial data, such as mortality rates together with scatter plots augmented with \emph{Pearsson's} $r$ values.
%%
A \emph{concept map} summarizes features related to a specified disease.
%%
Time-dependent epidemiological data are visualized by Chui et al.~\cite{Chui2011} using multi-panel graphs highlighting risk factor differences with age and gender with regard to influenza- and salmonellosis-associated hospitalizations.
%%
\magenta{
We incorporate the idea of the arrangement of feature combinations as matrix as well as assessing confounding features.
%%
}
%%
%We also include confounder features (age, gender) into the analysis while focusing on non-time-dependent cohort study data sets.
%%
\paragraph{Statistical analysis.}
%%
%Statistical analysis is an inherent part of Visual Analysis.
%%
%Here we present work showing a strong focus on statistical analysis in this context.
%% Quality Metrics in High-Dimensional Data Visualization - An Overview and Systematization
Bertini et al.~\cite{Bertini} present an overview of quality metrics describing high dimensional-data.
%%
Their research agenda acts as a guideline for our design.
%%
Namely, we apply aspects of perceptual tuning with human pattern recognition of important aspects, scalability between different data set sizes and application testing with domain experts.
%% The Sparse Regression Cube: A Reliable Modeling Technique for Open Cyber-physical Systems
Ahmadi et al.~\cite{Ahmadi} define the \emph{Sparse Regression Cube} that partitions sparse high dimensional data into subspaces, which are then described by their most reliable linear regression model.
%%
%In their application, they try to find most fuel-efficient roads connecting two user-defined landmarks.
%%
They focus on an algebraic representation for efficient regression-model calculation to find the best fit for a subspace.
%%
%The dimensions for the calculations are fixed, while we assess different features for each model calculation and can therefore not benefit from a hierarchical calculation approach.

@@

We extended the feature selection step (recall Sec. \ref{sec:Background}) with an iterative analysis loop incorporating group selection and visualization using expert input as well as clustering methods.

@@

The difference of the presented approach in comparison with our previous and related work is twofold.

@@

As described in Section~\ref{sec:RegressionAnalysis}, regression analysis is the statistical tool of choice for analyzing relationships in epidemiological data.
%%
A regression model is based on expert knowledge; there is no singular rule how to apply models to a given set of features. Thus, they have to be applied with care.
%%
% \design{
% %%
% Key requirements based on the epidemiological application domain on the 3D Regression Heat Map are:
% %%
% \begin{itemize}
% 	\item Show relationships of features w.r.t. their explanatory power towards the target feature (e.g., a disease or unhealthy condition).
% 	\item Allow for modeling assumptions regarding the target feature.
% 	\item The tool has to be easily accessible and simultaneously secure the data its processing.
% \end{itemize}
% %%
% }
\subsection{Regression Heat Map Description Using Regression Formula Notation}
%%
\design{The key requirement of modeling expert knowledge} is integrated into the analysis using \emph{regression formulas}.
%Expert knowledge is integrated into the analysis using regression formulas.
%%
As described in Section~\ref{sec:RegressionAnalysis}, the formula input influences the type of the chosen regression method as well as the \emph{independent} features describing the target.
%%

Since we want to associate the regression analyses with an overview visualization, we are interested in all possible combinations of (two or more) independent features describing a target.
We achieve this by introducing dynamic variables $X$, $Y$ and $Z$ into the regression notation.
Our method replaces the dynamic variables with all features in the data set.
In a data set with $n$ (e.g., 100) features, the regression formula $Cancer \sim X + Y$
%\begin{equation}
%
%\end{equation}
would yield $n^2$ (10,000) regression models, describing all possible combinations of two features in the data describing the target feature $Cancer$.
This notation comes natural to anyone familiar with regression analysis, since it is the standard way of expression.
With simple adjustments to the formula, different results can be achieved:
\begin{itemize}
	\item $Z \sim X + Y$ calculates all combinations of two features w.r.t. all possible target features.
	\item $Cancer \sim X + Y + BodyWeight$ includes the $BodyWeight$ feature into all regression models as feature with $Cancer$ as target.
	\item $Cancer \sim X + Y + Z$ calculates all combinations of three features w.r.t. the $Cancer$ target.
\end{itemize}
The problem with this approach lies in its complexity.
The number of calculated regression models exponentially increases for each dynamic variable added.
If we assume a data set with 100 features and the formula $Z \sim X + Y$, we obtain 1,000,000 regression models.

>>>>

%\com{New Solution}.
%%
We propose an analysis approach of epidemiological data sets by incorporating all features in an exhaustive regression-based analysis.
%%
This approach combines all \emph{independent features} w.r.t. a \emph{target feature}.
%%
It provides a visualization that reveals insights into the data by highlighting relationships.
%%
The \emph{3D Regression Heat Map}, a novel 3D visual encoding, acts as an overview of the whole data set.
%%
It shows all combinations of two to three independent features with a specific target disease.
%%
%A 3D visualization of all combinations of two to three independent features towards a target acts as an overview of the whole data set, the \emph{3D Regression Heat Map}.
%%
Slicing through the \emph{3D Regression Heat Map} allows for the detailed analysis of the underlying relationships.
%%
Expert knowledge about disease-specific hypotheses can be included into the analysis by adjusting the regression model formulas.
%%
Furthermore, the influences of features can be assessed using a difference view comparing different calculation results.
%%
%\com{Validation \& Results}.
%%
We applied our \emph{3D Regression Heat Map} method to a hepatic steatosis data set to reproduce results from a data mining-driven analysis.
%%
A qualitative analysis was conducted on a breast density data set.
%%
We were able to derive new hypotheses about relations between breast density density and breast lesions with breast cancer.
%%
%\com{Implications}.
%%
With the \emph{3D Regression Heat Map}, we present a visual overview of epidemiological data that allows for the first time an interactive regression-based analysis of large feature sets with respect to a disease.
} % end of abstract

@@

Testing features for associations with diseases using regression models is one of the most important epidemiological tools.
%%
Using regression analysis to assess the statistical resilience of a hypothesis rarely involves \emph{more than three features} due to the higher dimensional problem and the required subject count.
%%
Due to the amount of data and only limited overview visualizations, possible correlations may be missed.
%%
Explorative analyses and overview visualizations of the data set as presented in prior work \cite{Klemm2014VIS} are not tailored to a specific target feature.
%%
They mostly highlight correlations between features, which are known to the domain expert (e.g., correlation between body size and spine shape).
%%
We incorporate the regression analysis, which is familiar to the domain experts, into overview visualizations to support a hypothesis-free analysis or an analysis w.r.t. a specific disease or hypothesis.
%%
For this purpose we provide template regression formulas, which are applied to all potential feature combinations.
%%
Since the notation is familiar to epidemiologists, they can rapidly include their domain knowledge into the analysis process.
%%
Difference views between regression formulas allow to assess the influences of individual features on the process.
%%
Our contributions are:

@@

We described an Interactive Visual Analysis approach for image-centric cohort study data, which connects to the feature listing step \cite{Klemm2014VIS}.
%%
The methods aim to derive hypotheses through analysis of the data and observations about previously unknown feature correlations.
%%
\emph{Hypothesis generation} benefits from overview visualizations of feature correlations, which are not supported by standard statistical processors.
%%
In this work, we focus on a similar approach to derive insight and even new hypotheses through the data rather than only using it for a confirmatory analysis.
%%
\subsection{Epidemiological Data} \label{sec:EpidemiologicalData}
%%
There are many epidemiological instruments available to acquire data. %, which are suitable for different analysis types.
%%
We focus on population studies, which impose the largest data sets and yield a highly heterogenous and incomplete information space.
%%
Subject data are collected either regarding specific diseases or with the widest range possible.
%%
The latter allows the data set to be assessed w.r.t. different diseases.
%%
The feature space comprises information about lifestyle, somatometric features, medical parameters, genetic data as well as medical images.
%%
These features are gathered through questionnaires, medical examinations and laboratory analyses.
%%
Many features are sparse, such as follow-up questions about a medication or treatment of a certain disease.
%%
Other features are exclusive for a sub-group, such as women-specific questions, e.g., number of born children or period status.
%%
Medical status features or lifestyle factors are primarily of \emph{dichotomous} (binary) type.
%%
Continuous data are often discretized (e.g. 10 year steps for age) to equalize the feature types and to simplify the method selection.
%%
However, this reduces the information space and introduces a information bias, as assumptions are modeled through the discretization.
%%

Medical image data are also analyzed in modern population studies.
%%
We incorporate image-derived features, but do not focus on analyzing image data.
%%
More discussions of population study data types and characteristics can be found in the works of Preim et al.~\cite{Preim2014} and Toennies et al.~\cite{Toennies2015}.
%%

Features influencing the exposure as well as the outcome of an analysis are called \emph{confounders} and have to be specially treated.
%%
The analysis model has to be adjusted by normalizing all included features w.r.t. the confounder.
%%
\emph{Age} is included as confounder in almost any epidemiological analysis, since most diseases are more likely for older subjects.
%% TODO Shorten next sentence if needed
It also influences the general body condition and thereby almost all features acquired through population studies.
%%
Another important confounder is \emph{gender}.
%%
Confounders have to be selected by epidemiologists specific to the investigated condition.
%%
\paragraph{The Study of Health in Pomerania (SHIP).}
The \texttt{SHIP}, located in Northern Germany, aims to characterize health and disease in the widest range possible \cite{Volzke2011}.
%%
%It does not focus on a specific disease, making the data set open for many diseases.
%%
Unique for the \texttt{SHIP} is the acquisition of medical image data.
%%
A second cohort, \texttt{SHIP-TREND} was started in 2012.
%%
Data for both cohorts are examined in a 5-year time span.
%%
New parameters are added in each iteration, extending the range of investigated diseases.
%%
%For the latest two acquisitions (\texttt{SHIP-2} and \texttt{SHIP-TREND-0}), MRI scans are included into the cohort \cite{Hegenscheid2009, Ivanovska2014}.
%%
Most examinations occur in all stages and are performed according to the same instructions to enable comparisons over the different stages of the study.

\subsection{Regression Analysis} \label{sec:RegressionAnalysis}
%%
Regression analysis is the most important statistical tool when analyzing epidemiological data and are the basis of this work.
%%
A regression analysis assesses the influence of one or more (\emph{independent}) features to one target (\emph{dependent}) feature.
%%
The regression model yields a function describing the target feature by weighting the independent features.
%%
Different metrics, such as the weightings itself and associated $p$ values, describe the resulting function (the \emph{model}).
%%
\add{
%%
The \emph{Akaike Information Criterion} (AIC) metric estimates the quality of a model and is suited for comparing models \cite{Akaike}.
%%
A small AIC value indicates a higher quality model.
%%
}
$R^2$ values describe the quality of the fit; in other words how well the dependent features describe the target feature.
%%
The value ranges between $[0, 1]$, where 1 encodes a perfect fit.
%%

@@

\paragraph{Logistic Regression for Dichotomous Target.} Logistic regression implies a dichotomous target variable.
%%
The target is described by fitting a logistic function.
%%
Logistic models, as opposed to linear models, do not allow for extracting an $R^2$ quality of fit value.
%%
Therefore, pseudo-$R^2$ values are extracted, such as the \emph{Nagelkerke $R^2$} \cite{Nagelkerke}, which mimics the behavior of the $R^2$.
%%
\emph{Nagelkerke $R^2$} behaves different than $R^2$ values extracted from the linear regression model.
%%
Comparisons have to be handled with care.
%%
\section{Prior and Related Work}
%% https://en.wikipedia.org/wiki/Exploratory_data_analysis#cite_note-Tukey1977-4
Tukey already stated in 1977 that data are too often analyzed solely using confirmatory data analysis \cite{Tukey}.
%%
He emphasized the need to use data to \emph{derive} hypotheses, which can then be tested again.
%%
In this section we present prior and related work trying to achieve this goal.
%%
%The work of M\"{u}hlbacher et al.~\cite{Muehlbacher}, Piringer et al.~\cite{Piringer}, Chan et al.~\cite{Chan} and Zhang et al.~\cite{Zhang2014} are closest related to ours regarding different aspects.
\paragraph{Parameter space analysis using regression models.}
%%
\magenta{
Sedlmair et al.~\cite{Sedlmair} present a conceptual taxonomy of parameter space analysis.
%%
%It includes a abstract application domain independent data flow model, navigation strategies for exploring parameter spaces as well a characterization of analysis tasks.
%%
%Using the taxonomy, approaches can be classified to find works with similar problems and derive new design ideas and evaluate ideas.
%The taxonomy allows to classify approaches to find works with similar problems to derive new design ideas and evaluate ideas.
%%
Based on their input parameter taxonomy, we are visualizing \emph{model parameters} based on \emph{environmental parameters} in a \emph{global-to-local} navigation strategy.
%%
We solve a \emph{fitting} task by aiming to find models well suited for describing the input data.
}

\magenta{
%% A Partition-based Framework for Building and Validating Regression Models
The approach of M\"{u}hlbacher et al.~\cite{Muehlbacher} is closest to ours.
%%
They provide a framework for qualitative analyses of relationships and ranking features for numerical target features with regression models.
%%
%Correlations between features with the target yield a ordered matrix-plot, where feature combinations are used to depict models of interest.
%A view for single-relationships with the target and single features as well as a matrix-plot for feature combinations are used to depict models of interest.
%%
%The visualization allows to assess different model complexities.
%%
Existing regression models can be validated and compared using 3D views and 2D slice views.
%%
%We employ the same steps by preprocessing features using correlations with the target and providing an overview over combinations.
%%
M\"{u}hlbacher et al. focus on a smaller number of features, which can be assessed in more detail, yielding a plot matrix view, while we cover more features by abstracting the models.
%%
%We also take categorical targets and regression formula inputs into account.
%%
}

@@

\paragraph{Visual analysis of epidemiological data.}
%%
The work of Zhang et al.~\cite{Gotz2014, Zhang2014} is closest to ours regarding the application of visual analysis of cohort study data.
%%
They present \emph{Cohort Analysis via Visual Analytics} (\emph{CAVA}), a framework that distinguishes three major elements of cohort study data analysis: \emph{cohort data} (and its manipulation using operations), \emph{views} and \emph{analytics}.
%%
They use the system to find longitudinal pathways for diseases on the basis of health records.
%%
%Sub-groups are automatically created by incorporating \emph{analytics} and through selection in expert-guided \emph{views}.
%%
%They do not go into the detail w.r.t the analytics used, they seem to not incorporate confounder analysis (other than dividing groups by gender) and the user study lacks concrete goals.
%%
%Their data differs strongly from ours, having multiple moments for each subject, yielding a approach focused on finding pathways of interest.
%%
We incorporate their requirements, which consist of \emph{flexible} and \emph{iterative analysis}. %, meaning that the system has to adapt to different requirements in different analysis steps.
%%
%We implement this by supporting both hypothesis-based and hypothesis-free analysis.
%%
Angelelli et al.~\cite{Angelelli} visualize image-derived an non-image data using cube data structures with focus on comparison and knowledge extraction.
%%
%Their proposed engine represents multiple heterogenous cohort study data sets as normalized data cubes.
%%
%Thus, data redundancy is avoided and runtime measure aggregation can be carried out when measures of the different cubes have to be combined.
%%
%This allows for saving memory space.
%%
%Their presented aggregation functions allow for a fast data calculation to map it using different linked visualizations, such as scatter plots, bar charts or histograms.
%%
They use \emph{Pearsson's} $r$ to characterize relationships with target features and employ list views and scatter plots to visualize and rank them.
%%
%We use regression models and associated $R^2$ values to depict relationships between a target feature and multiple independent features instead of one and employ the information into one global visualization.
%%
In contrast, we focus on a fast large scale correlation analysis incorporating many features to derive insights.
%%
%\paragraph{Visual Analysis in Public Health.}
%%--
% Shneiderman et al.~\cite{Shneiderman2013} highlight the necessity of interactive visualizations in healthcare data. %, e.g. using \emph{Health 2.0}, which is an umbrella for web-based health-informatics frameworks.
% %%
% %They divide Health 2.0 in three domains, \emph{Personal health information}, \emph{clinical health information}, and \emph{public health information}.
% %%
% %We focus on the latter.
% %%
% %Shneiderman and colleagues also present challenges,
% The presented challenges in this application require the systems to be sufficiently comprehensible and intuitive to fit into the short time span that clinicians have for analyzing data.
% %%
% Furthermore, the systems should provide comparative relationship visualization, characterization of similarity and presenting risks.
%%
%We use the challenges as design guidelines for our web-based prototype, focusing on functional clean designs to allow fast insight into relationships towards diseases.
%%

@@

%% Steenwijk: Integrated Visual Analysis for Heterogeneous Datasets in Cohort Studies
Steenwijk et al.~\cite{Steenwijk} focus on \emph{hypothesis-free} exploration of cohort data.
%%
They employ a framework consisting of feature extraction and visualization to derive dependencies between image- and non-image features.
%%
They incorporate linked views using scatter plots, bar charts, parallel coordinates as well as time plots to display and brush the data.
%%
%Medical image data can be incorporated to investigate potentially interesting individuals.
%%
%%We distinguish to this and other linked-view analysis based approaches by showing abstracted information derived from a preceding analyze-first step to highlight relationships rather than displaying the actual data using various information visualization techniques.
%%
The techniques are useful for selected features or a data set with a small number of features due to their increase in complexity with every additional parameter.
%%
%They are therefore suitable for \emph{small} epidemiological data sets.
%%
\magenta{
Maries et al.~\cite{GRACE} propose \texttt{GRACE}, a framework for visually exploring correlations of features with spatial and non-spatial geriatric data.
%%
Sparse Partial Least Squares Regression and Tikhonov Regularization are used to produce predictor subsets and quantify correlation strengths.
%%
A multi view system linked spatial with non-spatial data and correlation coefficients are visualized using Kiviat diagrams (similar to star plots), correlations are plotted using p-values in the other views.
%%
%Dendograms are used to select different brain regions.
Instead of applying several linked views, we focus on one view of abstracted data.
}

Generalized Pairs plots (\texttt{GPLOM'S}) extend the concept of scatter plot matrices by the pairwise depiction of heterogeneous data using type-combination-dependent visualizations \cite{GPLOMS, Francois2013}.
%%
Dai et al.~\cite{Dai2005} incorporate a \texttt{GPLOM}-like visualization using choropleth maps mapping spatial data, such as mortality rates together with scatter plots augmented with \emph{Pearsson's} $r$ values.
%%
A \emph{concept map} summarizes features related to a specified disease.
%%
Time-dependent epidemiological data are visualized by Chui et al.~\cite{Chui2011} using multi-panel graphs highlighting risk factor differences with age and gender with regard to influenza- and salmonellosis-associated hospitalizations.
%%
\magenta{
We incorporate the idea of the arrangement of feature combinations as matrix as well as assessing confounding features.
%%
}
%%
%We also include confounder features (age, gender) into the analysis while focusing on non-time-dependent cohort study data sets.
%%
\paragraph{Statistical analysis.}
%%
%Statistical analysis is an inherent part of Visual Analysis.
%%
%Here we present work showing a strong focus on statistical analysis in this context.
%% Quality Metrics in High-Dimensional Data Visualization - An Overview and Systematization
Bertini et al.~\cite{Bertini} present an overview of quality metrics describing high dimensional-data.
%%
Their refer to their research agenda in our design.
%%
Namely, we apply aspects of perceptual tuning with human pattern recognition of important aspects, scalability between different data set sizes and application testing with domain experts.
%% The Sparse Regression Cube: A Reliable Modeling Technique for Open Cyber-physical Systems
Ahmadi et al.~\cite{Ahmadi} define the \emph{Sparse Regression Cube} that partitions sparse high dimensional data into subspaces, which are then described by their most reliable linear regression model.
%%
%In their application, they try to find most fuel-efficient roads connecting two user-defined landmarks.
%%
They focus on an algebraic representation for efficient regression-model calculation to find the best fit for a subspace.
%%
%The dimensions for the calculations are fixed, while we assess different features for each model calculation and can therefore not benefit from a hierarchical calculation approach.

@@

We extended the feature selection step (recall Sec. \ref{sec:Background}) with an iterative analysis loop incorporating group selection using expert input as well as clustering methods.

@@

The difference of the presented approach in comparison with related work is twofold.

@@

As described in Section~\ref{sec:RegressionAnalysis}, regression analysis is the statistical tool of choice for analyzing relationships in epidemiological data.
%%
A regression model is based on expert knowledge.
%%
There is no rule how to apply models to a given set of features. Thus, they have to be applied with care.
%%
% \design{
% %%
% Key requirements based on the epidemiological application domain on the 3D Regression Heat Map are:
% %%
% \begin{itemize}
% 	\item Show relationships of features w.r.t. their explanatory power towards the target feature (e.g., a disease or unhealthy condition).
% 	\item Allow for modeling assumptions regarding the target feature.
% 	\item The tool has to be easily accessible and simultaneously secure the data its processing.
% \end{itemize}
% %%
% }
\subsection{Regression Heat Map Description Using Regression Formula Notation}
%%
\design{Expert knowledge modeling} is carried out using \emph{regression formulas}.
%Expert knowledge is integrated into the analysis using regression formulas.
%%
The formula input influences the type of the chosen regression method as well as the \emph{independent} features describing the target (recall Sect.~\ref{sec:RegressionAnalysis}).
%%
